---
title: "p8105_hw2_yl4928"
author: "Rachel"
date: "10/3/2021"
output: github_document
---

```{r, message=FALSE}
library(tidyverse)
library(readxl)
```

## Problem 1

# Read and clean the Mr. Trash Wheel sheet
# omit rows that do not include dumpster-specific data

```{r, message = FALSE}
mtw_df = 
  read_excel("./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", range = "A2:N408") %>% 
  janitor::clean_names() %>% 
  na.omit() %>% 
  mutate(sports_balls = round(sports_balls))
```

Read and clean precipitation data for 2018 and 2019. For each, omit rows without precipitation data and add a variable for year. Next, combine precipitation datasets and convert month to a character variable (the variable month.name is built into R and should be useful).


```{r, message = FALSE}
precipitation_2018_df = 
   read_excel(
     "./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
   sheet = "2018 Precipitation",
   range = "A2:B14") %>% 
janitor::clean_names() %>%
drop_na(total) %>%
mutate(year = 2018)
```


```{r, message = FALSE}
precipitation_2019_df = 
    read_excel(
      "./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
    sheet = "2019 Precipitation",
    range = "A2:B14") %>% 
janitor::clean_names() %>% 
drop_na(total) %>% 
  mutate(year = 2019)
```
# combine precipitation datasets and convert month to a character variable

```{r, message = FALSE}
precipitation_df = 
  bind_rows(precipitation_2018_df, precipitation_2019_df) %>%
  janitor::clean_names() %>% 
  mutate(month = month.name[month])
```

Write a paragraph about these data

There are `r nrow(mtw_df)` observations in the Mr. Trash Wheel sheet. There are `r nrow(precipitation_df)` observations in the combined precipitation datasets of 2018 and 2019. Examples of key variables include `month` and `year`. The total precipitation in 2018 is `r precipitation_2018_df %>% pull(total) %>% sum`. The total precipitation in 2019 is `r precipitation_2019_df %>% pull(total) %>% sum`. The median number of sports balls in a dumpster in 2019 is `r mtw_df %>% filter(year == 2019) %>% pull(sports_balls) %>% median()`. 



## Question_2 

# part_a. 
clean the data in pols-month.csv. Use separate() to break up the variable mon into integer variables year, month, and day
replace month number with month name 
create a president variable taking values gop and dem, and remove prez_dem and prez_gop and remove the day variable. 

```{r, message = FALSE}
polsmonth_df = 
  read_csv("./data/fivethirtyeight_datasets/pols-month.csv") %>% 
  janitor::clean_names() %>% 
  separate(mon, into = c("year", "month", "day")) %>% 
  mutate(month = as.integer(month),
         month = month.name[month],
         president = ifelse(prez_gop == 0, "dem", "gop")) %>% 
  select(-prez_dem & -prez_gop & -day)
```



# Part_b
clean the data in snp.csv using a similar process to the above. For consistency across datasets, arrange according to year and month, and organize so that year and month are the leading columns.

```{r, message=FALSE}
 snp_df = 
  read_csv("./data/fivethirtyeight_datasets/snp.csv") %>% 
  janitor::clean_names() %>% 
  mutate(date = lubridate::mdy(date)) %>% 
  separate(date, into = c("year", "month", "day")) %>% 
  mutate(year = as.integer(year),
         month = as.integer(month),
         day = as.integer(day),
         month = month.name[month]
         ) %>% 
  arrange(year, month) %>% 
  relocate(year, month)
```




