---
title: "p8105_hw2_yl4928"
author: "Rachel"
date: "10/3/2021"
output: github_document
---

```{r}
library(tidyverse)
library(readxl)
```

## Problem 1

# Read and clean the Mr. Trash Wheel sheet
# omit rows that do not include dumpster-specific data

```{r}
mtw_df = 
  read_excel("./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", range = "A2:N408") %>% 
  janitor::clean_names() %>% 
  na.omit() %>% 
  mutate(sports_balls = round(sports_balls))
```

Read and clean precipitation data for 2018 and 2019. For each, omit rows without precipitation data and add a variable for year. Next, combine precipitation datasets and convert month to a character variable (the variable month.name is built into R and should be useful).


```{r}
precipitation_2018_df = 
   read_excel(
     "./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
   sheet = "2018 Precipitation",
   range = "A2:B14") %>% 
janitor::clean_names() %>%
drop_na(total) %>%
mutate(year = 2018)
```


```{r}
precipitation_2019_df = 
    read_excel(
      "./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx",
    sheet = "2019 Precipitation",
    range = "A2:B14") %>% 
janitor::clean_names() %>% 
drop_na(total) %>% 
  mutate(year = 2019)
```
# combine precipitation datasets and convert month to a character variable

```{r}
precipitation_df = 
  bind_rows(precipitation_2018_df, precipitation_2019_df) %>%
  janitor::clean_names() %>% 
  mutate(month = month.name[month])
```

Write a paragraph about these data

There are `r nrow(mtw_df)` observations in the Mr. Trash Wheel sheet. There are `r nrow(precipitation_df)` observations in the combined precipitation datasets of 2018 and 2019. Examples of key variables include `month` and `year`. The total precipitation in 2018 is `r precipitation_2018_df %>% pull(total) %>% sum`. The total precipitation in 2019 is `r precipitation_2019_df %>% pull(total) %>% sum`. The median number of sports balls in a dumpster in 2019 is `r mtw_df %>% filter(year == 2019) %>% pull(sports_balls) %>% median()`. 
